\thetitle

### 2021-03-03 writing/discussion.tex
discussion: start edits

There are a number of threats to the validity of these conclusions: (i) the
representativeness of benchmarks; (ii) the relation between mutations and
real programming mistakes; (iii) the definition of interesting debugging
scenarios; and (iv) the sampling strategy. The previous sections
have described the various ways we attempt to mitigate these threats;
the remainder of this section discussed three additional threats.

@@

Like {\em homo economicus\/}, which decouples the actual behavior of a
participant in an economy for the sake of mathematical modeling, the model of a
rational programmer decouples the actual debugging behavior of a software
developer for the sake of a systematic, large-scale analysis. This decoupling
comes with advantages and disadvantages. In the economic realm, mathematical models
have provided some predictive insights into the market's behavior; but as
behavioral economics has shown more recently, the mathematical abstraction of a
rational actor makes predictions also quite unreliable in some situations.
\footnote{It has also misled economists to focus on just the mathematics, though
this problem is not relevant here.}  Just like an ordinary consumer or producer,
an actual software developer is unlikely to stick to the exact strategy proposed
here. When this happens, the predicted benefits of blame assignment may not
materialize. Indeed, the authors's personal experience suggests such deviations, and
it also suggests that deviating is a mistake. To make a true judgment of the
usefulness of the rational-programmer idea, the community will need a lot more
experience with this form of evaluation and relating the evaluation to the
behavior of working programmers.

Relatedly, the experimental setup hides the type choice a rational programmer must make. When the
run-time checks signal an impedance mismatch in the real world, a programmer
does not have a typed component ready to swap in. Instead, the programmer must
come up with the next set of types---and this means the programmer must make
choices. It is normally possible to assign consistent types to variables in a
component in different ways. The creation and curation of the benchmarks over
many years has driven home this lesson, but fortunately, it has also shown that
the types are in most cases reasonably canonical.  The authors therefore conjecture that
a rational programmer would in most cases come up with an equivalent type
assignment in order to extend a trail as the experimental method describes.

>>>>

There are a number of threats to the validity of these conclusions: (i) the
representativeness of benchmarks; (ii) the relation between mutations and
real programming mistakes; (iii) the definition of interesting debugging
scenarios; and (iv) the sampling strategy. The previous sections
have described the various ways we attempt to mitigate these threats;
the remainder of this section discusses three additional threats.

@@

Like {\em homo economicus\/}, which decouples the actual behavior of a
participant in an economy for the sake of mathematical modeling, the model of a
rational programmer decouples the actual debugging behavior of a software
developer for the sake of a systematic, large-scale analysis. This decoupling
comes with advantages and disadvantages. In the economic realm, mathematical models
have provided some predictive insights into the market's behavior; but as
behavioral economics has shown more recently, the mathematical abstraction of a
rational actor makes predictions also quite unreliable in some situations.
\footnote{It has also misled economists to focus on just the mathematics, though
this problem is not relevant here.}  Just like an ordinary consumer or producer,
an actual software developer is unlikely to stick to the exact strategy proposed
here. When this happens, the predicted benefits of blame assignment may not
materialize. Indeed, the authors' personal experience suggests such deviations, and
it also suggests that deviating is a mistake. To make a true judgment of the
usefulness of the rational-programmer idea, the community will need much more
experience with this form of evaluation and relating the evaluation to the
behavior of working programmers.

Relatedly, the experimental setup hides the type choice a rational programmer must make. When the
run-time checks signal an impedance mismatch in the real world, a programmer
does not have a typed component ready to swap in. Instead, the programmer must
come up with the next set of types---and this means the programmer must make
choices. It is normally possible to assign consistent types to variables in a
component in different ways. The creation and curation of the benchmarks over
many years has driven home this lesson but, fortunately, it has also shown that
the types are in most cases reasonably canonical.  The authors therefore conjecture that
a rational programmer would in most cases come up with an equivalent type
assignment in order to extend a trail as the experimental method describes.
